Estimating Real-Time Highstreet Footfall from Wi-Fi Probe Requests

### 2018-06-30 paper/paper.tex
Almost completed methodology
Everything except main study is complete
Paul's comments included
end of day friday
start proofreading

\documentclass[]{interact}
\usepackage{epstopdf}
\usepackage{subfigure}
\usepackage{natbib}
\bibpunct[, ]{(}{)}{,}{a}{}{,}
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

@@

	\title{Estimating real-time highstreet footfall from Wi-Fi probe requests}
	\author{
		\name{ Balamurugan Soundararaj\textsuperscript{a}\thanks{CONTACT Balamurugan Soundararaj. Email: s.bala@ucl.ac.uk}, James Cheshire\textsuperscript{a} and Paul Longley\textsuperscript{a} }
		\affil{\textsuperscript{a}Department of Geography, University College London, United Kingdom.}
	}
	\maketitle

	\begin{abstract}
		\input{sections/abstract} \end{abstract}
	\begin{keywords}
		Highstreet footfall; Wi-Fi Probe requests; Sensors; MAC Randomisation \end{keywords}
	\section{Introduction}\label{introduction}
		\input{sections/introduction}
	\section{Previous Work}\label{previous_work}
		\input{sections/previous_work}
	\section{Methodology}\label{methodology}
		\input{sections/methodology}
	\section{Pilot Study}\label{pilot_study}
		\input{sections/pilot_study}
	\section{Main Study}\label{main_study}
		\input{sections/main_study}
	\section{Conclusion}\label{conclusion}
		\input{sections/conclusion}
	\section*{Acknowledgement}
		\input{sections/acknowledgement}
	\bibliographystyle{tfv} \bibliography{references}

>>>>

\documentclass[]{interact}
\usepackage{epstopdf}
\usepackage{subfigure}
\usepackage{natbib}
\bibpunct[, ]{(}{)}{,}{a}{}{,}
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}
\usepackage{color,soul}

@@

	\title{Estimating Real-Time Highstreet Footfall from Wi-Fi Probe Requests}
	\author{
		\name{ Balamurugan Soundararaj\textsuperscript{a}\thanks{CONTACT Balamurugan Soundararaj. Email: s.bala@ucl.ac.uk}, James Cheshire\textsuperscript{a} and Paul Longley\textsuperscript{a} }
		\affil{\textsuperscript{a}Department of Geography, University College London, United Kingdom.}
	}
	\maketitle
	\begin{abstract}
		\input{sections/abstract} \end{abstract}
	\begin{keywords}
		Pedestrian footfall; Urban sensing; Wi-Fi probe requests; MAC Randomisation \end{keywords}
	\section{Introduction}\label{introduction}
		\input{sections/introduction}
	\section{Background}\label{previous_work}
		\input{sections/previous_work}
	\section{Methodology}\label{methodology}
		\input{sections/methodology}
	\section{Pilot Study}\label{pilot_study}
		\input{sections/pilot_study}
	\section{Main Study}\label{main_study}
		\input{sections/main_study}
	\section{Conclusion}\label{conclusion}
		\input{sections/conclusion}
	\section*{Acknowledgement}
		\input{sections/acknowledgement}
	\bibliographystyle{tfv} \bibliography{references}

### 2018-06-30 paper/sections/abstract.tex
start proofreading

The accurate measurement human activity with high spatial and temporal granularity is crucial for us to understand the structure and function of built enviroment.
With increasing mobile ownerships, the Wi-Fi probe requests generated by these devices can be an excellent source of data for such measursement.
The major challanges in using Wi-Fi technology for such purpose are precisely delineating the field of measurement and arriving at accurate estimates without compromising the privacy of the users of the mobile devices.
In this paper we demonstrate that, with the application of class intervals and a novel graph based technique, we can overcome the challanges and reliably measure real-time footfall at retail highstreets in London.

>>>>

The accurate measurement of human activity with high spatial and temporal granularity is crucial for understanding the structure and function of the built environment.
With increasing mobile ownership, the Wi-Fi `probe requests' generated by mobile devices can act as a cheap, scalable and real-time source of data for establishing such measures.
The two major challanges we face in using these probe requests for estimating human activity are, filtering the noise generated by the uncertain field of measurement and clustering anonymised probe requests generated by the same devices together without compromising the privacy of the users.
In this paper we demonstrate that we can overcome these challanges by using class intervals and a novel graph based technique for filtering and clustering the probe requests which in turn, enables us to reliably measure real-time pedestrian footfall at retail highstreets.

### 2018-06-30 paper/sections/conclusion.tex
end of day friday
start proofreading

This paper has made a novel contribution to the measurement of activities in retail centres in real time.
To this end, it has described the collection and processing of a novel consumer Big Dataset that enables valid measures of levels of footfall activity which has been scaled across a Great Britain wide network of sensors.
In both conceptual and technical terms, it illustrates the ways in which passively collected consumer data can be ‘hardened’ to render them robust and reliable by using related procedures of internal and external validation.
Internal validation addresses the issues of screening out device probes that do not indicate footfall, and the further screening of device probes to ‘fingerprint’ the effects of MAC randomization.
External validation then entails reconciling adjusted counts with the footfall observed at sample locations.
The latter procedure makes it possible to generalise from locations at which manual footfall surveys are conducted to all others in the system, and to develop a classification of device locations that are more or less susceptible to noise generation.

It is important to note that the filtering process work based soley on the information present in the probe requests and their temporal distribution.
This ensures that although the mobile devices were uniquely identified, there was no further personal data generated by linking the probe requests to the users of the mobile devices.
This method essentially gave us a way to estimate the footfall in real-time without identifying or tracking the mobile devices themselves.

This Wi-Fi based footfall counting methodology offers a large number of applications and benefits for real time spatial analysis.
Since Wi-Fi based sensors are inexpensive and the data model is scalable, it is possible to use this methodology for a large network of sensors to gather granular data on pedestrian footfall.
A snapshot showing week's worth of precise footfall in area around charrig cross, London is shown in the \ref{} to demonstrate the potential for such dataset.
Projects such as SmartStreetSensors \citep{sss2016}, may utilise this methodology to overcome the challenges introduced by the implementation of MAC address randomisation.
Such precise and granular data also enables us to confidently model the pedestrian flow in urban road networks, and will be an indispensable tool in the smart city framework.
It can also be used to understand and classify geographical areas based on the spatio-temporal distribution of the volume of activity in them which we intend to research in future.

>>>>

This paper has made a novel contribution to the measurement of activities in retail centres in real time.
To this end, it has described the collection and processing of a novel consumer Big Dataset that enables valid measures of levels of footfall activity which has been scaled across a Great Britain wide network of sensors.
In both conceptual and technical terms, it illustrates the ways in which passively collected consumer data can be ‘hardened’ to render them robust and reliable by using related procedures of internal and external validation.

Internal validation addresses the issues of screening out device probes that do not indicate footfall, and the further screening of device probes to ‘fingerprint’ the effects of MAC randomization.It is important to note that the filtering process work based soley on the information present in the probe requests and their temporal distribution.
This ensures that although the mobile devices were uniquely identified, there was no further personal data generated by linking the probe requests to the users of the mobile devices.
This method essentially gave us a way to estimate the footfall in real-time without identifying or tracking the mobile devices themselves. 
External validation then entails reconciling adjusted counts with the footfall observed at sample locations.
This procedure makes it possible to generalise from locations at which manual footfall surveys are conducted to all others in the system, and to develop a classification of device locations that are more or less susceptible to noise generation.

This Wi-Fi based footfall counting methodology offers a large number of applications and benefits for real time spatial analysis.
Since Wi-Fi based sensors are inexpensive and the data model is scalable, it is possible to use this methodology for a large network of sensors to gather granular data on pedestrian footfall.
A snapshot showing week's worth of precise footfall in area around Charring cross, London is shown in Figure \ref{} to demonstrate the potential for such dataset.
Projects such as SmartStreetSensors \citep{sss2016}, may utilise this methodology to overcome the challenges introduced by the implementation of MAC address randomisation.
Such precise and granular data also enables us to confidently model the pedestrian flow in urban road networks, and will be an indispensable tool in the smart city framework.
It can also be used to understand and classify geographical areas based on the spatio-temporal distribution of the volume of activity in them which we intend to research in future.

### 2018-06-30 paper/sections/introduction.tex
start proofreading

There has been considerable research into the usefulness of these technologies to understand patterns of movement in cities in near real time \citep{candia2008uncovering, gonzalez2008understanding, calabrese2013understanding}. Most of this research has focused upon technical specification of accuracy or precision \citep{song2010limits, lane2010survey}, with somewhat less attention devoted to the ways that the characteristics of the technologies and of their human users conspire to create possible bias in the inclusiveness of usage patterns.
This is perhaps most manifest in the plethora of papers that analyse the distribution and content of publicly available Twitter social media feeds, which likely constitute an extremely biased and self-selecting subset of users that geo-enable their posts, drawn from a user base that itself bears no identified correspondence to any known background population \citep{lansley2016deriving}. 
Even analysis of mobile phone data, usually derived from industry players that have significant market share and user bases that may be representative of local populations, may exclude groups such as tourists from distand origins or subscribers to third party services that share distinctive characteristics \citep{di2016mind}
These examples illustrate the issues that underpin the assembly and analysis of consumer data, which can be considered as a distinctive class of Big Data that arise from the interactions between humans and customer-facing organisations such as retailers, domestic energy suppliers, transport providers and suppliers of social media and communications \citep{cdrc2018consumer}.
Consumer data account for an ever-increasing real share of all of the data that are collected of citizens, but a fundamental characteristic of consumer-led markets is that no single provider has a monopoly in market provision, and therefore issues of market share and segment generate bias in analysis.
The source and operation of this bias is unknown in the absence of extensive and context sensitive attempts to triangulate consumer data with data of known provenance relating to clearly defined populations \citep{lansley2016deriving}.
In similar ways to other classes of Big Data, consumer data are best throught of as digital ‘exhaust’, or a by-product created by or harvested from consumer transactions.

In this paper, we evaluate the value of data collected from a Great Britain wide network of 800 devices, installed in order to characterize the footfall patterns of a scientifically balanced sample of retail centres.
These devices are located in shop windows and record the probes emitted by mobile phones and other WiFi enabled devices.
These are deemed to be consumer data because devices carried by consumers routinely probe for a consumer service, specifically a WiFi connection.
Monitoring the probes from such devices provides an indication of the presence of their users, regardless of whether or not Internet connectivity is established.
Our core motivation is to appraise the usefulnessof WiFi probe records harvested from our network of sensors in order to indicate levels of pedestrian activity.
More broadly still, in our future research we intend to classify the nationwide network of footfall profiles as part of a programme of research to understand the form and functioning of retail areas at a time of far-reaching structural change for the retail industry.

To this end, it is important to first undertake a thorough conceptual and technical appraisal of our consumer data source.
In technical terms, screening of information present in the `probes' and classifing them based on their characteristics is essential in order to remove the ones emitted by devices that do not indicate pedestrian activity, such as network enabled printers and other fixed devices.
Related to this, a method to figerprint Wi-Fi probes is necessary to remove probes from individuals’ devices that in conceptual terms should not be considered part of footfall – as when, for example, an employees is seated in an office within range of the sensor device.
A calibration of sensor measurement is also essential on two grounds: first, individuals may carry multiple devices, or no device at all; and second, the positioning and orientation of the sensor in the retail unit may lead to systematic over- or under-enumeration.
These sources of bias in measurement must be accommodated by manual recording of footfall at each location and generalization of these sample survey results to all locations and time periods.
As we describe in detail below, manual validation of the data needs to be undertaken in parallel with technical profiling of the mix of consumer devices that probe our sensors, since the effectiveness of data cleaning procedures discussed in this paper differ between individual locations and configurations.

>>>>

There has been considerable research into the usefulness of these technologies to understand patterns of movement in cities in near real time \citep{candia2008uncovering, gonzalez2008understanding, calabrese2013understanding}. Most of this research has focused upon technical specification of accuracy or precision \citep{song2010limits, lane2010survey}, with somewhat less attention devoted to the ways that the characteristics of the technologies and of their human users conspire to create possible bias in the inclusiveness of usage patterns.
This is perhaps most manifest in the plethora of papers that analyse the distribution and content of publicly available Twitter social media feeds, which likely constitute an extremely biased and self-selecting subset of users that geo-enable their posts, drawn from a user base that itself bears no identified correspondence to any known background population \citep{lansley2016deriving}. 
Even analysis of mobile phone data, usually derived from industry players that have significant market share and user bases that may be representative of local populations, may exclude groups such as tourists from distand origins or subscribers to third party services that share distinctive characteristics \citep{di2016mind}

These examples illustrate the issues that underpin the assembly and analysis of consumer data, which can be considered as a distinctive class of Big Data that arise from the interactions between humans and customer-facing organisations such as retailers, domestic energy suppliers, transport providers and suppliers of social media and communications \citep{cdrc2018consumer}.
Consumer data account for an ever-increasing real share of all of the data that are collected of citizens, but a fundamental characteristic of consumer-led markets is that no single provider has a monopoly in market provision, and therefore issues of market share and segment generate bias in analysis.
The source and operation of this bias is unknown in the absence of extensive and context sensitive attempts to triangulate consumer data with data of known provenance relating to clearly defined populations \citep{lansley2016deriving}.
In similar ways to other classes of Big Data, consumer data are best throught of as digital ‘exhaust’, or a by-product created by or harvested from consumer transactions.

In this paper, through a set of experiments, we evaluate the value of data collected from a Great Britain wide network of 800 devices, installed in order to characterize the footfall patterns of a scientifically balanced sample of retail centres.
These devices are located in shop windows and record the probes emitted by mobile phones and other WiFi enabled devices.
These are deemed to be consumer data because devices carried by consumers routinely probe for a consumer service, specifically a WiFi connection.
Monitoring the probes from such devices provides an indication of the presence of their users, regardless of whether or not Internet connectivity is established.
Our core motivation is to appraise the usefulnessof WiFi probe records harvested from our network of sensors in order to indicate levels of pedestrian activity.
More broadly still, in our future research we intend to classify the nationwide network of footfall profiles as part of a programme of research to understand the form and functioning of retail areas at a time of far-reaching structural change for the retail industry.

To this end, it is important to first undertake a thorough conceptual and technical appraisal of our consumer data source.
In technical terms, screening of information present in the `probes' and classifing them based on their characteristics is essential in order to remove the ones emitted by devices that do not indicate pedestrian activity, such as network enabled printers and other fixed devices.
Related to this, a method to figerprint Wi-Fi probes is necessary to remove probes from individuals’ devices that in conceptual terms should not be considered part of footfall – as when, for example, an employees is seated in an office within range of the sensor device.
A calibration of sensor measurement is also essential on two grounds: first, individuals may carry multiple devices, or no device at all; and second, the positioning and orientation of the sensor in the retail unit may lead to systematic over- or under-enumeration.
These sources of bias in measurement must be accommodated by manual recording of footfall at each location and generalization of these sample survey results to all locations and time periods.
As we describe in detail below, manual validation of the data needs to be undertaken in parallel with technical profiling of the mix of consumer mobile devices that probe our sensors, since the effectiveness of data cleaning procedures discussed in this paper differ between individual locations and configurations.

### 2018-06-30 paper/sections/pilot_study.tex
start proofreading

To start we designed a small pilot study with the aim to validate the classification and clustering methodology against the scale and complexity of data collected on a open public area such as a retail highstreet.
We also need to find the algorithm which is best suited for classification of signal strengths to filter out the background noise.
The data was collected at Oxford Street, London on 20 December 2017 from 12:30 to 13:00 hrs, where Wi-Fi probe requests were collected using the sensor described in Chapter \ref{methodology} and pedestrian footfall was manually recorded using the Android app \citep{bala2018clicker}.
Being located at one of the busiest retail locations in the United Kingdom, the Wi-Fi sensor captured approximately 60,000 probe requests during the half hour period, and 3,722 people were recorded manually walking on the sidewalk during that time.

>>>>

To start we designed a small pilot study to validate the classification and clustering methodology against the scale and complexity of data collected on a open public area such as a retail highstreet.
We also aim to find the algorithm which is best suited for classification of signal strengths to filter out the background noise.
The data was collected at Oxford Street, London on 20 December 2017 from 12:30 to 13:00 hrs, where Wi-Fi probe requests were collected using the sensor described in Chapter \ref{methodology} and pedestrian footfall was manually recorded using the Android app \citep{bala2018clicker}.
Being located at one of the busiest retail locations in the United Kingdom, the Wi-Fi sensor captured approximately 60,000 probe requests during the half hour period, and 3,722 people were recorded manually walking on the sidewalk during that time.

### 2018-06-30 paper/sections/previous_work.tex
start proofreading

In the past decade Wi-Fi has emerged as the most commonly used technology in providing high speed internet access to mobile devices such as smartphones, tablets and laptops in public and private spaces. This has resulted in multiple Wi-Fi networks being available at almost every location in dense urban environments. Traversing through this overlapping mesh of Wi-Fi networks, modern mobile devices with Wi-Fi antennae regularly broadcast a special type of signal known as 'Probe Requests' in order to discover Wi-Fi networks available to them. This helps these devices to connect and switch between the WiFi networks seamlessly.

Probe requests are low level signals standardised by IEEE 802.1b/g specification \citep{ieee2013} as the first step in establishing a Wi-Fi based connection between two devices and is implemented in any Wi-Fi capable device irrespective of the manufacturer or the model. This ubiquity and standardisation make them an excellent source of open, passive, continuous, and wireless data generated by Wi-Fi capable devices present at any given time and location. Considering the unprecendented levels of mobile device ownership in recent years, we can in turn use this data to understand the population distribution in highly dynamic urban environments with high spatial and temporal granularity \citep{freud2015,konto2017}.

While a Wi-Fi based method to collect data offers us various advantages such as, easy scalability and efficiency in terms of cost and time, It also introduces few systematic biases, uncertainities in the collected data along with the serious risk of infringing on the privacy of the mobile users. In this paper, using a set of probe requests and manual counts collected at various high street locations across London, we demonstrate that pedestrian footfall at these locations can be estimated with considerable precision and accuracy while protecting the privacy of the pedestrians.

Though WiFi is a `location-less' technology, there are reliable methods to triangulate the location of Wi-Fi enabled mobile devices by the known locations of APs and the signal strength reported by them from the mobile devices \citep{he2003range, moore2004robust, lamarca2005place}.
This can overcome the usual shortcoming of GPS, which struggles for precision and accuracy in indoor and densely built environments \citep{zarim2006,kawaguchi2009wifi, xi2010locating}. 
Utilising this, we can easily and quickly estimate trajectories of the mobile devices just using the WiFi communication the device has with multiple known APs \citep{Sorensenlocation} which can be used similar to the GPS trajectories to understand individual travel patterns (Kim, 2006;\citep{reki2007,Sap2015}, crowd behaviour \citep{abedi2013bluetooth,mowafi2013tracking}, vehicular \citep{lu2010vehicle} and pedestrian movement \citep{xu2013pedestrian,fukuzaki2014pedestrian,wang2016gait}.
Such data can also be used in transportation planning and management to estimate travel time \citep{musa2011wiflow} and real time traffic monitoring \citep{abbott2013empirical}.

@@

Though extensive research has been carried out on this subject with feasible and favorable results, in recent years, one of the major challenges faced in such attempts has been the increasing attempt by mobile phone manufacturers to protect the users’ privacy by anonymising the globally identifiable portion of the probe requests, \citep{green2008}.
There are various methods which have been devised to overcome this anonymisation process such as decomposition of OUIs where detailed device model information is estimated by analysing an already known dataset of OUIs \citep{martin2016decomposition}; Scrambler attack using a small part of the physical layer specification for WiFi \citep{vo2016,bloessl2015scrambler}; and finally, the timing attack where the packet sequence information present in the probe request frame is used \citep{matte2016,cheng2016can}.
A combination of these methodologies has been proven to produce de-anonymised globally unique device information \citep{vanhoef2016, martin2017}. These approaches usually result in serious risk of infringement of the privacy of the users of the mobile devices by revealing their identifiable personal information.

There is a clear gap in the research for exploring methodologies which enable us to estimate the number of unique mobile devices from a set of anonymised probe requests, without the need to reveal their original MAC addresses.
Such technique has various applications in numerous fields such as uncovering the urban wireless landscape \citep{rose2010mapping}, revealing human activity at large scales \citep{qin2013discovering}, estimating pedestrian numbers in crowds \citep{schauer2014estimating,fukuzaki2015statistical} and even counting people in hyper local scales such as queues \citep{wang2013measuring}
With enough infrastructure to collect such information we can aim to generate a real-time census of the city \citep{konto2017}.
It has also been demonstrated by \citep{pinelli2015comparing} through series of experiments on a telecom operator dataset that using such network-driven approach is more advantageous compared to the widely used event-driven approaches. With this background we set out to device and implement a methodology to reliably estimate human activity such as pedestrian footfall from Wi-Fi probe requests without infringing the privacy of the users involve.

>>>>

In the past decade Wi-Fi has emerged as one of the most commonly used technology in providing high speed internet access to mobile devices such as smartphones, tablets and laptops in public and private spaces.
This has resulted in multiple Wi-Fi networks being available at almost every location in dense urban environments. 
Traversing through this overlapping mesh of Wi-Fi networks, modern mobile devices with Wi-Fi antennae regularly broadcast a special type of signal known as `Probe Requests' in order to discover Wi-Fi networks available to them.
This helps these devices to connect and switch between the Wi-Fi networks seamlessly.

Probe requests are low level signals standardised by IEEE 802.1b/g specification \citep{ieee2013} as the first step in establishing a Wi-Fi based connection between two devices and is implemented in any Wi-Fi capable device irrespective of the manufacturer or the model.
This ubiquity and standardisation make them an excellent source of open, passive, continuous, and wireless data generated by Wi-Fi capable devices present at any given time and location.
Considering the unprecendented levels of mobile device ownership in recent years, we can in turn use this data to understand the population distribution in highly dynamic urban environments with high spatial and temporal granularity \citep{freud2015,konto2017}. While a Wi-Fi based method to collect data offers us various advantages such as, easy scalability and efficiency in terms of cost and time, It also introduces few systematic biases, uncertainities in the collected data along with the serious risk of infringing on the privacy of the mobile users.
In this paper, using a set of probe requests and manual counts collected at various high street locations across London, we demonstrate that pedestrian footfall at these locations can be estimated with considerable precision and accuracy while protecting the privacy of the pedestrians.

Though Wi-Fi is a `location-less' technology, there are reliable methods to triangulate the location of Wi-Fi enabled mobile devices by the known locations of APs and the signal strength reported by them from the mobile devices \citep{he2003range, moore2004robust, lamarca2005place}.
This can overcome the usual shortcoming of GPS, which struggles for precision and accuracy in indoor and densely built environments \citep{zarim2006,kawaguchi2009wifi, xi2010locating}. 
Utilising this, we can easily and quickly estimate trajectories of the mobile devices just using the Wi-Fi communication the device has with multiple known APs \citep{Sorensenlocation} which can be used similar to the GPS trajectories to understand individual travel patterns (Kim, 2006;\citep{reki2007,Sap2015}, crowd behaviour \citep{abedi2013bluetooth,mowafi2013tracking}, vehicular \citep{lu2010vehicle} and pedestrian movement \citep{xu2013pedestrian,fukuzaki2014pedestrian,wang2016gait}.
Such data can also be used in transportation planning and management to estimate travel time \citep{musa2011wiflow} and real time traffic monitoring \citep{abbott2013empirical}.

@@

Though extensive research has been carried out on this subject with feasible and favorable results, in recent years, one of the major challenges faced in such attempts has been the increasing attempt by mobile phone manufacturers to protect the users’ privacy by anonymising the globally identifiable portion of the probe requests, \citep{green2008}.
There are various methods which have been devised to overcome this anonymisation process such as decomposition of OUIs where detailed device model information is estimated by analysing an already known dataset of OUIs \citep{martin2016decomposition}; Scrambler attack using a small part of the physical layer specification for Wi-Fi \citep{vo2016,bloessl2015scrambler}; and finally, the timing attack where the packet sequence information present in the probe request frame is used \citep{matte2016,cheng2016can}.
A combination of these methodologies has been proven to produce de-anonymised globally unique device information \citep{vanhoef2016, martin2017}.
These approaches usually result in serious risk of infringement of the privacy of the users of the mobile devices by revealing their identifiable personal information.

There is a clear gap in the research for exploring methodologies which enable us to estimate the number of unique mobile devices from a set of anonymised probe requests, without the need to reveal their original MAC addresses.
Such technique has various applications in numerous fields such as uncovering the urban wireless landscape \citep{rose2010mapping}, revealing human activity at large scales \citep{qin2013discovering}, estimating pedestrian numbers in crowds \citep{schauer2014estimating,fukuzaki2015statistical} and even counting people in hyper local scales such as queues \citep{wang2013measuring}
With enough infrastructure to collect such information we can aim to generate a real-time census of the city \citep{konto2017}.
It has also been demonstrated by \citep{pinelli2015comparing} through series of experiments on a telecom operator dataset that using such network-driven approach is more advantageous compared to the widely used event-driven approaches.
With this background we set out to device and implement a methodology to reliably estimate human activity such as pedestrian footfall from Wi-Fi probe requests without infringing the privacy of the users involved.
